---
title: "Assignment 1 - Language Development in ASD - part 2"
author: "Riccardo Fusaroli"
date: "July 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Language development in Autism Spectrum Disorder (ASD)

Background: Autism Spectrum Disorder is often related to language impairment. However, this phenomenon has not been empirically traced in detail: i) relying on actual naturalistic language production, ii) over extended periods of time. We therefore videotaped circa 30 kids with ASD and circa 30 comparison kids (matched by linguistic performance at visit 1) for ca. 30 minutes of naturalistic interactions with a parent. We repeated the data collection 6 times per kid, with 4 months between each visit. We transcribed the data and counted: 
i) the amount of words that each kid uses in each video. Same for the parent.
ii) the amount of unique words that each kid uses in each video. Same for the parent.
iii) the amount of morphemes per utterance (Mean Length of Utterance) displayed by each child in each video. Same for the parent. 

This data is in the file you prepared in the previous class. 

NB. A few children have been excluded from your datasets. We will be using them next week to evaluate how good your models are in assessing the linguistic development in new participants.

We then want to test the language trajectory of child and parent over time.

This Markdown is structured in the following way:

1. The exercises: read them carefully. Under each exercise you will have to write your answers, once you have written and run the code. This is the part that you have to directly send to the teachers.
2. An (optional) guided template full of hints for writing the code to solve the exercises. Fill in the code and the paragraphs as required. Then report your results under the exercise part.
3. In exercise 4 you will be asked to create the best possible model of language development in TD and ASD children, picking and choosing whatever additional variables you want from the dataset. Next time, the models produced by the different groups will compete against each other to see who can produce the best model, so choose carefully!

You will have to have a github repository for the code and send the answers to Celine and Riccardo without code (but a link to your github repository). This way we can check your code, but you are also forced to figure out how to report your analyses :-)

N.B. The following lines are a summary of the questions to be answered, the step-by-step instructions and tips are below.

## Exercise 1) Preliminary Data Exploration

Describe the participant samples in the dataset (e.g. by diagnosis, age, etc.). Do you think the two groups are well balanced? If not, what do you think was the reason?

TD children are significantly younger than ASD children, by approx. 12.62 months. There is no significant difference in the genders across the groups. There is a significant difference between TD and ASD children in ADOS scores, TD children have scores on average 13.37 points lower than ASD children.There is a significant difference between ASd and TD, where TD children have an average score of 2.78 point higher with the verbal IQ. TD children have a significantly higher use of words than ASD, on average 182 words.TD have on average 52.09 more unique words than ASD children which is significantly different.TD has a significantly longer MLU by 0.66 than ASD. TD parents have significantly longer MLU than ASD parents, by 0.49 on average.

Due to the major age difference between the TD and the ASD children, I do not find the gruops balanced. In a young age, children develope quickly, in which case it does not make sense to compare a 2 year old  to a 3 year old with ASD. However it is curious how the significantly younger TD children er better than the ASD. 

### Exercise 2) Children learning language: the effects of time and ASD
Describe linguistic development in TD and ASD children in terms of Mean Length of Utterance (MLU)?

The childrens MLU changes significantly over time, but is however not significantly affecetd by diagnosis. ( Beta = 0.23, se = 0.02, t-value = 9.516, p<0.05). The R^2m that explains the models fixed effects (Visit + Diagnosis) explains 22% of the varience. 

### Exercise 3) Child directed speech as a moving target
Describe how parental use of language changes over time in terms of MLU. What do you think is going on?

MLU of a parent changes significantly over time (Beta=0.12, se= 0.02, t value= 6.54, p<0.05). Furthermore the MLU of a parent is also is also dependent of ia a child has a ASD diagnosis. Parents of a child TD has a significantly longer MLU than perent of a ASD child (Beta=0.50, se= 0.12, t value=4.36, p<0.05)

It may be due to that the parent obseve the child and therefore if the child has ASD and thereby have a short MLU, the parent will automatically shorten their MLU aswell so the child can keep up. 

### Exercise 4) Looking into "individual differences" (demographic, clinical or cognitive profiles)
The dataset contains some additional variables characterizing the kids’ cognitive and clinical profile: ADOS (autism severity), MSEL EL (Expressive Language, that is, verbal IQ, or linguistic skills at first visit as assessed by a psychologist using Mullen Scales of Early Learning), MSEL VR (Visual Reception, used as a proxy for non verbal IQ at first visit), Age, Gender, Ethnicity. Would it make sense to add any of them to your model of linguistic trajectories? Create the best possible model (the one that best explain the data, with MLU as outcome). Next time your model will be tested on new participants, and we will proclaim a winner. Describe your strategy to select the best models (how did you choose the variables to include?) and send the code to Riccardo and Celine.

By looking at the R2m we can see that fixed effects of model 5 (m1) explains 56% of the varience compared to the 37% explained by model 4. The chosen varibales were chosen by the significanse shown by the anova tests.


### [OPTIONAL] Exercise 5) Comment on how the three linguistic variables measure linguistic performance (the so-called "construct validity" of the measures). Do they express the same variance?

[HERE GOES YOUR ANSWER]

### Structure of the code chunks

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format?) and describe the participants (by diagnosis)

We will try to answer three questions:

- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?
  
### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries}
library(dplyr)
library(ggplot2) 
library(lme4)
library(pastecs)


```

### Define your working directory and load the data

- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

```{r Load Data}

setwd("~/Assignment-1.1")
data= read.csv("AutismData1.csv")

```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r}
# Comparing TD's and ASD's (use lm for t.test)
data$ID = as.factor(data$ID)
data$Visit = as.factor(data$Visit)
# Stats on Age
data1 = subset(data, Visit == 1)
by(data1$Age, data1$Diagnosis, stat.desc)
ageTest = lm(Age ~ Diagnosis, data = data1)
summary(ageTest)
#plots 
ggplot(data1, aes(Diagnosis, Age)) + geom_boxplot()
ggplot(data1, aes(Diagnosis, Age)) + geom_violin()
# TD children are significantly younger than ASD children, by approx. 12.62 months 
# Gender 
genderStats = glm(Gender ~ Diagnosis, family = binomial, data = data)
summary(genderStats)
## ggplot(data, aes(Gender, Diagnosis)) + geom_boxplot()
# No significant difference in gender across the two groups, p > 0.262. 
# ADOS test 
adosStats = lm(ADOS ~ Diagnosis, data = data) 
summary(adosStats)
# There is a significant difference between TD and ASD children in ADOS scores, TD children have scores on average 13.37 points lower than ASD children. 
# Verbal IQ 
viqStats = lm(verbalIQ ~ Diagnosis, data = data) 
summary(viqStats)
## There is a significant difference between ASd and TD, where TD children have an average score of 2.78 point higher with the verbal IQ. 
# Non_verbal IQ 
nviqStats = lm(nonVerbalIQ ~ Diagnosis, data = data) 
summary(nviqStats)
## There is an insignificant difference in Non verbal IQ between TD and ASD children, however TD children show a slightly lower (0.89) non verbal IQ than ASD children. 
# Number of words used 
wuStats = lm( tokens_CHI ~ Diagnosis, data = data) 
summary(wuStats)
## TD children have a significantly higher use of words than ASD, on average 182 words.
# Number of unique words used 
uwuStats = lm( types_CHI ~ Diagnosis, data = data) 
summary(uwuStats)
## Significant different, TD have on average 52.09 more unique words than ASD children.
# MLU child
mluChildStats = lm( CHI_MLU ~ Diagnosis, data = data) 
summary(mluChildStats)
## TD significantly longer MLU by 0.66 than ASD. 
#MLU parent 
mluMotStats = lm( MOT_MLU ~ Diagnosis, data = data)
summary(mluMotStats)
## TD parents have significantly longer MLU than ASD parents, by 0.49 on average. 
```

[REPORT THE RESULTS]

## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r}
data$Gender = as.factor(data$Gender)
data$Gender = recode(data$Gender, "1"="F", "2"="M")
data$Visit=as.numeric(data$Visit)
## Hypothesis 1 
data$Diagnosis = as.factor(data$Diagnosis)
data$Diagnosis = revalue(data$Diagnosis, c("1"="ASD", "2"="TD"))
## we know 1 = ASD and 2 = TD, so change back to numeric so we can run linear model on it
data$Diagnosis = as.numeric(data$Diagnosis)



library(lmerTest)
install.packages("lmerTest")

# testing 
diagnosisTest = lmer(CHI_MLU ~ Visit + Diagnosis + (1 + Visit|ID), data = data, REML = FALSE)
summary(diagnosisTest)


mluHypothesis = lmer(CHI_MLU ~ Diagnosis  + (1 + Visit|ID), data = data,REML=FALSE)
summary(mluHypothesis)


anova(diagnosisTest,mluHypothesis)

#The childrens MLU changes significantly over time, but is however not significantly affecetd by diagnosis. ( Beta = 0.23, se = 0.02, t-value = 9.516, p<0.05). 

```

How would you evaluate whether the model is a good model?

```{r}
r^2

install.packages("MuMIn")
library(MuMIn)

r.squaredGLMM(diagnosisTest)

#The R^2m explains the models fixed effects (Visit + Diagnosis) explains 22% of the varience. 

```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better

```{r}

```

Exciting right?
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results

[REPORT THE RESULTS]
Linguistic development of children MLU is affected by ... [COMPLETE]

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r}
## Hypothesis MOT MLU 
mluMotHypothesis = lmer(MOT_MLU ~ Visit + Diagnosis + (1 + Visit|ID), data = data)
summary(mluMotHypothesis)



##MLU of a parent changes significantly over time (Beta=0.12, se= 0.02, t value= 6.54, p<0.05). Furthermore the MLU of a parent is also is also dependent of ia a child has a ASD diagnosis. Parents of a child TD has a significantly longer MLU than parents of a ASD child (Beta=0.50, se= 0.12, t value=4.36, p<0.05)
```

[REPORT THE RESULTS]

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Riccardo and Celine.


```{r}
cltModel1 = lmer(CHI_MLU ~  Visit+Ethnicity + (1 + Visit|ID), data = data)
summary(cltModel1)
cltModel2 = lmer(CHI_MLU ~ Visit + nonVerbalIQ + (1 + Visit|ID), data = data)
summary(cltModel2)
cltModel3 = lmer(CHI_MLU ~ Visit+ verbalIQ + (1 + Visit|ID), data = data)
summary(cltModel3)
cltModel4 = lmer(CHI_MLU ~ Visit + MOT_MLU+nonVerbalIQ + (1 + Visit|ID), data = data)
summary(cltModel4)
m1 = lmer(CHI_MLU ~ Visit + MOT_MLU + verbalIQ + (1+Visit|ID), data=data, REML =F)
summary(m1)

## run Anova 
anova(cltModel1, cltModel2, cltModel3, cltModel4,m1)


r.squaredGLMM(m1)
r.squaredGLMM(cltModel4)

#By looking at the R2m we can see that fixed effects of model 5 (m1) explains 56% of the varience compared to the 37% explained by model 4. 

```

REPORT OF THE RESULTS