---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the third exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time, then learning how to cross-validate models and finally how to systematically compare models.

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/ra99bdvm6fzay3g/demo_test.csv?dl=0
* Utterance Length data: https://www.dropbox.com/s/uxtqqzl18nwxowq/LU_test.csv?dl=0
* Word data: https://www.dropbox.com/s/1ces4hv8kh0stov/token_test.csv?dl=0

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data () and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the code again and apply it to Assignment2TrainData1.csv)
- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the functions rmse() and predict() )
- create the test dataset (apply the code from assignment 1 part 1 to clean up the 3 test datasets)
- test the performance of the models on the test data (Tips: time to reuse "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())


In order to calculate our model's performances on the training data and the test data, we look the root mean sqare error, which is a desciption of how far the datapoints are from our general linear model. The RSME for training data is 0.34. When applying the model to the test data, the RSME raises to 0.70 - which still is not bad. In conclusion, the model is doing good in describing the training data and the test data.  


```{r setup, include=FALSE}
# Load library
library(dplyr)
library(ggplot2) 
library(lme4)
library(pastecs)

# Set working directory
setwd("~/Assignment-1.1")

# Load data
LU_test= read.csv("LU_test.csv")
token_test=read.csv("Token_test.csv")
demo_test=read.csv("demo_test.csv")


#Clean data
names(LU_test)[names(LU_test)=="VISIT"]="Visit"
names(token_test)[names(token_test)=="VISIT"]="Visit"

names(token_test)[names(token_test)=="SUBJ"]="ID"
names(demo_test)[names(demo_test)=="Child.ID"]="ID"
names(LU_test)[names(LU_test)=="SUBJ"]="ID"

library(stringr)
LU_test$Visit=str_extract(LU_test$Visit, "\\d")
token_test$Visit=str_extract(token_test$Visit, "\\d")

LU_test$ID=gsub("\\.", "", LU_test$ID)
demo_test$ID=gsub("\\.", "", demo_test$ID)
token_test$ID=gsub("\\.", "", token_test$ID)
library(dplyr)

demo_test = select(demo_test, ID, Visit, Ethnicity, Diagnosis, Gender, Age, ADOS, MullenRaw, ExpressiveLangRaw)
LU_test = select(LU_test, ID, Visit, MOT_MLU, MOT_LUstd, CHI_MLU, CHI_LUstd)
token_test = select(token_test, ID, Visit, types_MOT, types_CHI, tokens_MOT, tokens_CHI)

names(demo_test)[names(demo_test)=="MullenRaw"]="nonVerbalIQ"
names(demo_test)[names(demo_test)=="ExpressiveLangRaw"]="verbalIQ"

data1=merge(LU_test,token_test,by=c("ID","Visit"))
  
data=merge(data1,demo_test,by=c("ID","Visit")) 

visit1 = subset(data[data$Visit == "1",])

visit1=select(visit1,ADOS,verbalIQ,nonVerbalIQ,ID)

data=data[-15:-17]

data=merge(visit1,data,by=c("ID"))
#Rename ID to numbers

data$ID=as.factor(data$ID)
levels(data$ID)=1:length(levels(data$ID))

#
data$Visit=as.numeric(data$Visit)
levels(data$Visit)=1:length(levels(data$Visit))

#Rename gender
library(plyr)
data$Gender=as.factor(data$Gender)
data$Gender = revalue(data$Gender, c("1"="M", "2"="F"))

data$Diagnosis= revalue(data$Diagnosis,c("A"="ASD","B"="TD"))

write.csv(data,file="AutismData_test.csv")


data_old=read.csv("AutismData1.csv")
data_new=read.csv("AutismData_test.csv")

# Our Code
library(lmerTest)

m1 = lmer(CHI_MLU ~ Visit + MOT_MLU + verbalIQ + (1+Visit|ID), data_old, REML =F)

install.packages("modelr")
library(modelr)

rmse(m1,data_new)


# In order to calculate the models performance, we look the root mean sqare error. The RSME for training data is 0.34. When applying the model to the test data, the RSME raises to 0.70 - which still is not bad. 

#Predict???

```

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Create the basic model of ChildMLU as a function of Time and Diagnosis (don't forget the random effects!).
- Make a cross-validated version of the model. (Tips: google the function "createFolds";  loop through each fold, train a model on the other folds and test it on the fold)
- Report the results and comment on them.

- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: How would you go comparing the performance of the basic model and the cross-validated model on the testing set?
- Bonus Question 2: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 3: compare the cross-validated predictive error against the actual predictive error on the test data

For this exercise a loop was created in order to crossvalidate to calculate the predictive error. By doing so we saw how our original model's fixed effects only could explain 23% of the varience. Then we created 4 other models including other predictors. By running them throug the loop we could see one of the new models could explain 55% of the varience. 

```{r setup, include=FALSE}
#Model

m2 = lmer(CHI_MLU ~ Visit + Diagnosis + (1+Visit|ID), data_new, REML = F)

modelr::rmse(m2, data_new)
r.squaredGLMM(m2)
# Crossvalidation

install.packages("groupdata2")
install.packages("Metrics")
library(groupdata2)
library(MuMIn)


trainingdata= fold(data_old, k=4, cat_col = "Diagnosis", id_col = "ID")

#Creating function for automatic crossvalidation. Outputs R2c, R2m and RMSE for each fold, as well the mean values across folds
cv = function(data, k, model, dependent){
#Creating variables for storing performances
rmselist = list()
r2list = list()
#Creating loop
for (i in 1:k){
  train = data[data$.folds != i,]    #creating training set (all folds except the one)
  validation = data[data$.folds == i,] #creating testing/validation set (the current fold)
  model = lmer(model, train, REML = F)   #running lmer on the model specified in the function call
  rmselist[i] = Metrics::rmse(validation[[dependent]], predict(model, validation, allow.new.levels = T))  #saving model rmse
  r2list[i] = as.data.frame(r.squaredGLMM(model))     #saving r2c and r2m
}
#doing some wrangling so the R2 outputs can be printed in a nice format
r2list = as.data.frame(t(as.data.frame(r2list)))
colnames(r2list) = c("R2m", "R2c")
rownames(r2list) = seq(1:k)
r2list = as.data.frame(r2list)

#returning the wanted values
return(c('RMSE' = rmselist, 'Mean RMSE' = mean(unlist(rmselist)), r2list,  'Mean R2m' = mean(r2list$R2m), 'Mean R2c' =  mean(r2list$R2c)))
}

m1 = "CHI_MLU ~ Visit + Diagnosis + (1+Visit|ID)"
cv(trainingdata, 4, m1, 'CHI_MLU')

lmer("CHI_MLU ~ Visit + Diagnosis + (1+Visit|ID)")



#Our models fixed effects can only predict 23% of the varience. 

#Now we try new models

m2 = "CHI_MLU ~ Visit + Diagnosis +MOT_MLU + (1+Visit|ID)"
cv(trainingdata, 4, m2, 'CHI_MLU')

#30% of the varience

m3 = "CHI_MLU ~ Visit + Diagnosis+verbalIQ + (1+Visit|ID)"
cv(trainingdata, 4, m3, 'CHI_MLU')
#51% of the varience

m4 = "CHI_MLU ~ Visit + Diagnosis+ MOT_MLU+verbalIQ+ (1+Visit|ID)"
cv(trainingdata, 4, m4, 'CHI_MLU')
#55% of the varience
 





```

### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.
(Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; input the average of the TD group  for each parameter in the model as X1, X2, etc.).

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

Bernie perform better compared to the avarage TD child. Bernie has by visit 1 a longer MLU of 0.67. At visit 2 he has longer MLU of 0.78. At visit 3 he has longer MLU of 1.12. At visit 4 he has longer MLU of 0.44. At visit 5 he has longer MLU of 0.20. At visit 6 he has longer MLU of 0.52.

Our model can predict how a childs MLU would developed based on the mothers MLU and the verbal IQ of the child. Bernie performs close to how the model predicts. At visit 6 Bernie has a MLU of 3.45. The model predicts Bernie to have a MLU of 3.62 - So our model miss with 0.18 - which is not a lot. 


```{r}

#crl+alt+i for ny rmarkdown


bernie=subset(data_new,ID=="2")

model=lmer(CHI_MLU ~ Visit + Diagnosis+ MOT_MLU+verbalIQ+ (1+Visit|ID),data_old)
summary(model)

tsub=subset(data_old,Diagnosis=="TD")

avgMOTMLU = tsub %>% 
  group_by(Visit) %>% 
  dplyr::summarize(MOT_MLU = mean(MOT_MLU))

avgCHI_MLU = tsub %>%
  group_by(Visit) %>%
  dplyr::summarize(CHI_MLU = mean(CHI_MLU))

avgCHI_MLU

td_chi = data.frame(ID = rep(300, 6), Visit = seq(1,6), Diagnosis="TD",verbalIQ=mean(tsub$verbalIQ),MOT_MLU=avgMOTMLU[2], CHI_MLU = avgCHI_MLU[2])

bernie=dplyr::select(bernie,ID,Visit,Diagnosis,verbalIQ,MOT_MLU,CHI_MLU)
bernie

avgTD_bernie = rbind(td_chi, bernie)
avgTD_bernie$Visit = as.factor(avgTD_bernie$Visit)
avgTD_bernie$ID = as.factor(avgTD_bernie$ID)

avgTD_bernie

install.packages("ggplot2")
library(ggplot2)

#Lille flot plot
ggplot(avgTD_bernie, aes(Visit, CHI_MLU, color = ID)) +
  geom_point()+ 
  geom_line()

#Shows how much better Bernie is a every visit compared to the avarage TD child.
bernie$CHI_MLU-td_chi$CHI_MLU

#Predict Bernies CHI_MLU at visit 6
bernie_predict=predict(model, bernie, allow.new.levels = T)

bernie$CHI_MLU-bernie_predict

bernie_predict

#Bernie perform worse at the 6th visit compared to the average TD child by 0.18

```



### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).

Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.

- Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?

### OPTIONAL: Exercise 5): Using Lasso for model selection
Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!


